### 2024-08-24 layout/format_booklet.tex
Add figures

\usepackage[paperheight=240mm,paperwidth=160mm]{geometry}

\usepackage{float}
\floatplacement{figure}{p}
\floatplacement{table}{p}