### 2024-09-16 chapters/preface.tex
BCI chapter

\chapter{Preface}
The mind works in mysterious ways\ldots\
Let's try to hack it!

I'd like to acknowledge