### 2024-09-16 layout/format_booklet.tex

\usepackage[paperheight=240mm,paperwidth=160mm]{geometry}

\usepackage{float}
\floatplacement{figure}{p}
\floatplacement{table}{p}

\usepackage{incgraph}
\usepackage{tikz}
%\newcommand{\fullpagefig}[3]{
%  \incgraph[
%    overlay={
%    }
%  ]{#1}
%}
\definecolor{lightgray}{HTML}{ECECEC}

\usepackage[usenames,x11names,dvipsnames,svgnames]{xcolor}
\usetikzlibrary{backgrounds}
\newcommand{\fullpagefig}[3]{%
  \igrcenter{\igrboxset{\includegraphics{#1}}%
  \begin{tikzpicture}[background rectangle/.style={fill=lightgray}, show background rectangle]]%
  \igrboxtikzcenter%
  %\draw[help lines] (0,0) grid (\igrBoxWidth,\igrBoxHeight);
  \node[above right,text width=10cm] at ([xshift=1cm,yshift=1cm]page.south west){{\captionof{figure}{#2}}};
  \end{tikzpicture}}
  \label{#3}
}