### 2024-09-16 chapters/abstract.tex
Add table, references for eye motor disabilities

Main contribution: improve gaze-independent decoding by isolating component
related to visuospatial attention from visual components