### 2024-09-20 main.tex

% ATTENTION: Don't build this, either build main_booklet.tex or
% main_manuscript.tex
%
% Preamble ====================================================================
\pdfminorversion=7
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[inline]{enumitem}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Commands
\input{math.tex}
\input{acronyms.tex}
\newcommand{\includechapter}[2]{
  \chapter{#1}
  \begin{refsection}
  \input{#2}
    \printbibliography[heading=subbibliography]
  \end{refsection}
  \clearpage
}

\newcommand{\comment}[1]{
  \todo[color=yellow]{#1}
}

% Bibliography
\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}
\addbibresource{publications.bib}

% Title

\begin{document}

% Front matter ================================================================
\frontmatter

\include{chapters/preface.tex}
\chapter{Abstract}
\include{chapters/abstract.tex}
\chapter{Samenvatting}
\tableofcontents
\listoffigures
\listoftables
\printacronyms[name=List of acronyms]

% Main matter =================================================================
\mainmatter%

\part{Introduction \& Literature}
\includechapter{Brain-Computer Interfaces}{chapters/bci.tex}
\includechapter{Gaze-independent visual BCIs}{chapters/gaze_independence.tex}
\todo{The role of jitter in gaze-independent ERP decoding}
%\includechapter{Problem statement \& rationale}{chapters/problem_statement.tex}

\part{Algorithms \& Decoders}
\includechapter{Kronecker-structured linear decoding}{chapters/stbf_struct.tex}

\chapter{Tensor discriminant analysis}
\todo{add Tensorizations coping with latency jitter}

\includechapter{Classfier-based Latency Estimation with Woody iterations}{chapters/wcble.tex}

\part{Experiments \& Applications}
\includechapter{Compensating jitter for gaze-independent decoding}{chapters/covert_align.tex}
\todo{Integrating gaze-tracking data}
\includechapter{Patient cases}{chapters/patients.tex}
\includechapter{Conclusion \& recommendations}{chapters/conclusion.tex}

\backmatter%
\chapter{Funding \& acknowledgments}
\chapter{Personal contribution}
\chapter{Conflict of interest}
The authors declare no conflict of interest.
\include{chapters/cv.tex}
\end{document}