### 2024-09-24 chapters/abstract.tex
Intro, P300 stimulation

\todo{Main contribution: improve gaze-independent decoding by isolating component
related to visuospatial attention from visual components}