### 2024-09-24 main.tex
Set manuscript draft

% ATTENTION: Don't build this, either build main_booklet.tex or
% main_manuscript.tex
%
% Preamble ====================================================================
\pdfminorversion=7
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[inline]{enumitem}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Commands
\input{math.tex}
\input{acronyms.tex}
\newcommand{\includechapter}[2]{
  \chapter{#1}
  \begin{refsection}
  \input{#2}
    \printbibliography[heading=subbibliography]
  \end{refsection}
  \clearpage
}

\newcommand{\comment}[1]{
  \todo[color=yellow]{#1}
}
\newcommand*\elide{\textup{[\,\dots]}}

% Bibliography
\usepackage[backend=biber, doi=false, url=false, isbn=false, eprint=false]{biblatex}
\addbibresource{references.bib}
\addbibresource{publications.bib}

% Set bibliography font size
\renewcommand*{\bibfont}{\footnotesize}

\usepackage[inline]{enumitem}

\begin{document}

% Front matter ================================================================
\frontmatter

\include{chapters/preface.tex}
\chapter{Abstract}
\include{chapters/abstract.tex}
\chapter{Samenvatting}
\tableofcontents
\listoffigures
\listoftables
\printacronyms[name=List of acronyms]

% Main matter =================================================================
\mainmatter%

\part{Introduction \& Literature}
\includechapter{Brain-computer interfaces}{chapters/bci.tex}
\includechapter{Gaze-independent visual BCIs}{chapters/gaze_independence.tex}
\todo{The role of jitter in gaze-independent ERP decoding}
%\includechapter{Problem statement \& rationale}{chapters/problem_statement.tex}

\part{Algorithms \& Decoders}
\includechapter{Kronecker-structured linear decoding}{chapters/stbf_struct.tex}

\chapter{Tensor discriminant analysis}
\todo{insert tensor paper}
\todo{discuss Tensorizations coping with latency jitter to make link}

\includechapter{Classfier-based Latency Estimation with Woody iterations}{chapters/wcble.tex}

\part{Experiments \& Applications}
\includechapter{Compensating jitter for gaze-independce}{chapters/covert_align.tex}
\todo{Integrating gaze-tracking data}
\includechapter{Patient cases}{chapters/patients.tex}
\includechapter{Conclusion \& recommendations}{chapters/conclusion.tex}

\backmatter%
\chapter{Funding \& acknowledgments}
\todo{my grant}\todo{marc's projects}
\chapter{Personal contribution}
\todo{Juliette, Alixe, Kristl Claeys, Edward Schietekatte, Marc, François,
Hakim, Marie,Etienne, Philip Van Damme  }
\chapter{Conflict of interest}
The authors declare no conflict of interest.
\include{chapters/cv.tex}
\end{document}