### 2024-09-25 figures/bci/paradigms.tex
Refactor structure

\documentclass[tikz,border=10pt]{standalone}
\usepackage{fontspec}

\begin{document}
\begin{tikzpicture}

    % Define custom colors
    \definecolor{darkgray}{RGB}{38,35,33}
    \definecolor{lightgray}{RGB}{143,130,125}
    \definecolor{accent1}{RGB}{242,144,24} % Formerly orange
    \definecolor{accent2}{RGB}{246,61,30}  % Formerly red
    \definecolor{accent3}{RGB}{120,71,157}  % Formerly purple
    \definecolor{white}{RGB}{255,255,255}

    % Draw the horizontal line
    \draw[darkgray, thick, ->] (36, 120) -- (124, 120);

    % Draw the vertical line
    \draw[darkgray, thick, ->] (80, 76) -- (80, 164);

    % Draw rectangles
    \draw[accent1, thick] (39, 79) rectangle (73.5, 113.5);
    \draw[accent2, thick] (86, 79) rectangle (120.5, 113.5);
    \draw[accent3, thick] (39, 125) rectangle (120, 158);

    % Add text
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{white}] at (56.5, 90) {};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{lightgray}] at (64.7, 70) {active};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{lightgray}] at (74, 172) {passive};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{lightgray}] at (19, 118) {stimulus\\independent};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{lightgray}] at (139, 118) {stimulus\\dependent};

    % Text in rectangles
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (109, 136) {error\\detection};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (97, 93) {P300};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (56, 90) {motor\\imagery};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (56, 101) {imagined\\speech};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (96, 103) {SSVEP};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (110, 93) {cVEP};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (109, 104) {mVEP};

    % More text
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (56, 133) {attention\\monitoring};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (56, 146) {emotion\\monitoring};
    \node[align=center, font=\fontsize{8}{10}\sffamily\color{darkgray}] at (90, 146) {media induced\\emotion recognition};

\end{tikzpicture}
\end{document}