### 2024-09-25 include/layout/figures_booklet.tex

\usepackage{incgraph}
\usepackage{tikz}

\usetikzlibrary{backgrounds}

\renewcommand{\fullpagefig}[3]{%
  \igrcenter{\igrboxset{\includegraphics{#1}}%
  \begin{tikzpicture}[background rectangle/.style={fill=lightgray}, show background rectangle]]%
  \igrboxtikzcenter%
  %\draw[help lines] (0,0) grid (\igrBoxWidth,\igrBoxHeight);
  \node[above right,text width=10cm,align=justify] at ([xshift=1cm,yshift=1cm]page.south west){{\captionof{figure}{#2}}};
  \end{tikzpicture}}
  \label{#3}
}
\renewcommand{\fullpagefigct}[3]{%
  \igrcenter{\igrboxset{\includegraphics{#1}}%
  \begin{tikzpicture}[background rectangle/.style={fill=lightgray}, show background rectangle]]%
  \igrboxtikzcenter%
  %\draw[help lines] (0,0) grid (\igrBoxWidth,\igrBoxHeight);
  \node[below right,text width=10cm,align=justify] at ([xshift=1cm,yshift=-1cm]page.north west){{\captionof{figure}{#2}}};
  \end{tikzpicture}}
  \label{#3}
}

\renewcommand{\figpane}[3]{
    \begin{subfigure}{.45\textwidth}
        \includegraphics[width=\textwidth]{#1}
        \caption{#2}
        \label{#3}
    \end{subfigure}
}

% TODO nice spacing and background
\renewcommand{\fourpanefig}[6]{
    \begin{figure}
      #1\hfill#2

      #3\hfill#4

    \caption{#5}
    \label{#6}
    \end{figure}
}