### 2024-10-02 chapters/patients_table.tex

\begin{tabular}{llllllllr}
  \toprule
  \textbf{ID}  & \textbf{Age} & \textbf{Sex} & \textbf{Hand.} &
  \textbf{Diagnosis}
  & \textbf{Speech}     & \textbf{Trach.} & \textbf{Communication}          &
  \textbf{Cls.} \\ \midrule
  PA1 & 58  & M   & L     & bulbar-onset \acs{als} & anarthric  & no          & tablet                 & 3  \\
  PB1 & 41  & M   & L     & \acs{fa} & dysarthric & no          & verbal                 & 3  \\
  PB2 & 43  & F   & R     & \acs{fa} & dysarthric & no          & verbal                 & 3  \\
  PB4 & 48  & M   & R     & \acs{fa} & dysarthric & no          & verbal                 & 3  \\
  PC2 & 43  & M   & R     & brainstem stroke & anarthric  & yes         & \makecell[l]{prompting\\+eye movement} & 2 \\
  PC3 & 43  & F   & R     & brainstem stroke & anarthric  & yes         & letterboard            & 2 \\
  PC4 & 54  & M   & R     & \makecell[l]{left cerebellar stroke \\ (trombosis of the basilar artery)} & anarthric  & yes & letterboard & 2 \\
  \bottomrule
\end{tabular}