### 2024-10-02 figures/wcble/smearing.tikz.tex

\begin{tikzpicture}
    % First subplot (low jitter) positioned relative to center
    \begin{axis}[
        at={(-3cm,0cm)}, % Adjust position to the left, closer to the center
        anchor=center,
        width=7cm, height=6cm,
        xmin=20, xmax=80,
        ymin=-0.1, ymax=1.1,
        axis lines=none, % Remove axes
        title={Low Jitter}, % Add title
        title style={yshift=-10pt}, % Adjust title position to make it closer to the plot
        ]
        % Plot individual waveforms (low jitter) in darkgray
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-52)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-48)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-51)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-49)/5)^2)};

        % Plot average waveform (low jitter) in accent1 color, very thick
        \addplot[ultra thick,accent1,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};
    \end{axis}

    % Second subplot (high jitter) positioned relative to center
    \begin{axis}[
        at={(3cm,0cm)}, % Adjust position to the right, closer to the center
        anchor=center,
        width=7cm, height=6cm,
        xmin=20, xmax=80,
        ymin=-0.1, ymax=1.1,
        axis lines=none, % Remove axes
        title={High Jitter}, % Add title
        title style={yshift=-10pt}, % Adjust title position to make it closer to the plot
        ]
        % Plot individual waveforms (high jitter) in darkgray
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-45)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-55)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-40)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-60)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};

        % Plot average waveform (high jitter) in accent1 color, very thick
        \addplot[ultra thick,accent1,domain=20:80,samples=100] {0.2*(exp(-0.5*((x-45)/5)^2) + exp(-0.5*((x-55)/5)^2) + exp(-0.5*((x-40)/5)^2) + exp(-0.5*((x-60)/5)^2) + exp(-0.5*((x-50)/5)^2))};
    \end{axis}

    % Unified legend positioned further below the center
    \node at (0,-5) {\begin{tikzpicture}
        \begin{axis}[
            hide axis,
            xmin=0, xmax=0, ymin=0, ymax=0,
            legend style={
	    	draw=none, legend columns=2, /tikz/every even column/.append
		style={column sep=2cm}
		}
            ]
            \addlegendimage{darkgray,thick}
            \addlegendentry{Waveforms}
            \addlegendimage{accent1,ultra thick}
            \addlegendentry{Average}
        \end{axis}
    \end{tikzpicture}};
\end{tikzpicture}
\vspace{-5cm}