### 2024-10-03 figures/gaze_independence/lis_venn.tikz.tex

\sffamily
\begin{tikzpicture}

    % Draw the outer lightgray box with ultra thick darkgray line
    \draw[fill=lightgray, draw=darkgray, ultra thick] (-4,-1.5) rectangle (4,5.5);
    % Label for the outer box in the top left corner
    \node[anchor=north west, text=darkgray] at (-4,5.5) {\textbf{\acs{bci} users}};

    % Draw the outermost circle for SSPI with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent1] (0,2) circle (3);
    \node[text=accent1] at (0,4) {\textbf{\acs{sspi}}};  % Centered bold text with line color

    % Draw the middle circle for SSPGI with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent2] (0,1) circle (2);
    \node[text=accent2] at (0,1.66) {\textbf{\acs{sspgi}}};  % Centered bold text with line color

    % Draw the innermost circle for LIS with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent3] (0,0) circle (1);
    \node[text=accent3] at (0,0) {\textbf{\acs{lis}}};  % Centered bold text with line color

\end{tikzpicture}