### 2024-10-04 figures/bci/erp.tikz.tex

\sffamily
\sansmath
\footnotesize

\begin{tikzpicture}[trim axis left]
    \begin{axis}[
        width=\textwidth,
        height=0.6180\textwidth,
        scale only axis,
        xlabel={Time (ms)},
        ylabel={Potential ($\mu$V)},
        axis lines=middle,
	ylabel near ticks,
	xlabel near ticks,
	y label style={darkgray},
	x label style={darkgray},
        ymin=-5, ymax=10,
        xmin=-100, xmax=800,
        enlarge x limits=false,
        enlarge y limits=false,
        smooth,
        every axis plot/.append style={very thick, accent1}, % Set the line color to accent1
        ytick={-5,0,5}, % Y-ticks at -5, 0, 5
        xtick={200,400,600}, % X-ticks at 200, 400, 600, 800
        axis line style={darkgray}, % Set the axis line color to darkgray
        tick style={darkgray}, % Set the tick color to darkgray
        tick label style={darkgray} % Set the tick label color to darkgray
    ]
    % ERP waveform using cubic splines for smooth connection
    \addplot[
        smooth,
        tension=0.5
    ] coordinates {
        (-100, 0)  % Starting point
        (0, 0)     % Baseline
        (50, 0)    % Pre-P1
        (80, 2)    % P1
        (115, -3)  % N1
        (160, 3)   % P2
        (220, 1)   % N2
        (400, 7)   % P3
        (650, 1)   % LPP (Late Positive Potential)
        (800, 1)   % End point
    };

    % Annotations for the components
    \node[anchor=south, color=muteblack] at (axis cs:80,2) {P1};
    \node[anchor=north, color=muteblack] at (axis cs:115,-3) {N1};
    \node[anchor=south, color=muteblack] at (axis cs:160,3) {P2};
    \node[anchor=west, color=muteblack] at (axis cs:220,1) {N2};
    \node[anchor=south, color=muteblack] at (axis cs:400,7) {P3};
    \node[anchor=south west, color=muteblack] at (axis cs:650,1) {LPP};

    \end{axis}
\end{tikzpicture}