### 2024-10-04 figures/bci/paradigms.tikz.tex

\sffamily
\begin{tikzpicture}[
    scale=\textwidth/2cm,
  ]
  \useasboundingbox (-1,-1.25) rectangle (1,1);
  \pgfmathsetmacro{\margin}{.1}
  \pgfmathsetmacro{\center}{.5+.5*\margin}
  \pgfmathsetmacro{\textspacing}{.15}
  %\begin{scope}[shift={(1,1.2)}]
    % Draw axes
    \draw[darkgray, very thick, <->] ({-(1+\margin)}, 0) -- ({1+\margin},0);
    \draw[darkgray, very thick, <->] (0,{-(1+\margin)}) -- (0,{1+\margin});

    % Draw rectangles
    \draw[draw=accent1, fill=white, very thick] ({-\margin}, \margin) rectangle (-1,1);
    \draw[draw=accent2, fill=white, very thick] (\margin, \margin) rectangle (1,1);
    \draw[draw=accent3, fill=white, very thick] (-1, {-\margin}) rectangle (1,-1);

    %% Add text
    \node[color=accent1, align=left,font=\bfseries, anchor=north west] at  (-1,1){active};
    \node[color=accent2, align=left,font=\bfseries, anchor=north east] at  (1,1){reactive};
    \node[color=accent3, align=left,font=\bfseries, anchor=south west] at  (-1,-1){passive};

    \node[color=darkgray, align=center,font=\bfseries, anchor=north] at
    (0,{-(1+\margin)}) {passive\\participation};
    \node[color=darkgray, align=center,font=\bfseries, anchor=south] at
    (0,{1+\margin}) {active\\participation};
    \node[color=darkgray, align=center,font=\bfseries, anchor=east] at ({-(1+\margin)}, 0) {stimulus\\independent};
    \node[color=darkgray, align=center,font=\bfseries, anchor=west] at ({1+\margin},0) {stimulus\\dependent};

    %% Text in rectangles
    \node[align=center] at ({-\center}, {\center+\textspacing}) {motor\\imagery};
    \node[align=center] at ({-\center}, {\center-\textspacing}) {imagined\\speech};

    \node[align=center] at ({\center+\textspacing}, {\center+\textspacing}) {P300};
    \node[align=center] at ({\center+\textspacing}, {\center-\textspacing}) {SSVEP};
    \node[align=center] at ({\center-\textspacing}, {\center+\textspacing}) {cVEP};
    \node[align=center] at ({\center-\textspacing}, {\center-\textspacing}) {mVEP};

    %% More text
    \node[align=center] at ({\center-\textspacing}, {-\center+\textspacing})  {error\\potentials};
    \node[align=center] at ({-\center-\textspacing}, {-\center+\textspacing}) {attention\\monitoring};
    \node[align=center] at ({-\center+\textspacing}, {-\center-\textspacing}) {emotion\\monitoring};
    \node[align=center] at ({\center-\textspacing}, {-\center-\textspacing}) {neurofeedback};
  %\end{scope}
\end{tikzpicture}