### 2024-10-04 fig.tikz.tex

\documentclass{standalone}
\usepackage{dissertation}
\input{chapters/bttda/tensorviz.tex}

\begin{document}
\begin{tikzpicture}[y=-1cm]
    \draw[help lines] (-3,-3) grid (20,6);
    \fill[red] (0,0) circle (2pt);

	  \begin{scope}[shift={(-0.0707, -0.0707)}]
      \MatrixSkewed{$\mat{A}_1$}{$r_1$}{$D_1$}{2}{1}
    \end{scope}
	  \begin{scope}[shift={(-0.1,0)}]
      \MatrixLeft{$\mat{A}_2$}{$r_2$}{$D_2$}{2}{1}
    \end{scope}
	  \begin{scope}[shift={(0,1.1)}]
      \MatrixBelow{$\mat{A}_3$}{$r_3$}{$D_3$}{1}{2}
    \end{scope}
    \TensorThree{$\ten{G}$}{}{}{}{1}{1}{1}

    \node at (2,.5){$=$};

	  \begin{scope}[shift={(3,-.5)}]
    \TensorThree{$\ten{X}^\prime$}{$D_1$}{$D_2$}{$D_3$}{2}{2}{2}
    \end{scope}

    \node at (6.5,.5){$=$};

	  \begin{scope}[shift={(7.5,-.5)}]
    \TensorThree{$\ten{X}$}{$D_1$}{$D_2$}{$D_3$}{2}{2}{2}
    \end{scope}

    \node at (11,.5){$-$};

	  \begin{scope}[shift={(12,-.5)}]
      \TensorThree{$\ten{\mathbfcal{E}}$}{$D_1$}{$D_2$}{$D_3$}{2}{2}{2}
    \end{scope}

\end{tikzpicture}
\end{document}