### 2024-10-06 chapters/conflict.tex
cover page

\chapter*{Conflict of interest}
The authors declare no conflict of interest.