### 2024-10-06 chapters/contribution.tex

\chapter*{Personal contribution}
\todo{which works conceptualized}
\todo{Juliette, Alixe, Kristl Claeys, Edward Schietekatte, Marc, François,
Hakim, Marie,Etienne, Philip Van Damme  }