### 2024-10-06 chapters/funding.tex

\chapter*{Funding \& acknowledgments}
\todo{my grant}\todo{marc's projects}

\todo{acknowledgement wcble paper}
We thank the Flemish Supercomputer Center (VSC) and the High-Performance
Computing (HPC) center of the KU Leuven for allowing us to execute our
computational experiments on their systems.
We also wish to acknowledge dr.\ Axel Faes for his inspiration in conceptualizing
the work presented in Chapter~\ref{sec:bttda}.

\todo{people at TRAINM, MAS and Kristl Claeys}
\todo{juliette meunier}