### 2024-10-06 cover_kul.tex

\newgeometry{margin=1.5cm}
\begin{titlepage}

    \sffamily
    \begin{minipage}[t]{.3\textwidth}
    \vspace{0pt}
    KU Leuven \\
    Biomedical Sciences Group \\
    Faculty of Medical Science \\
    Department of Neurosciences

    \end{minipage}\hfill%
    \begin{minipage}[t]{.3\textwidth}
      \vspace{0pt}
      \includegraphics[width=\textwidth]{figures/kul_docschool.png}
      \bigskip

      \includegraphics[width=\textwidth]{figures/ulille.png}
    \end{minipage}

    \vfill

    \noindent\makebox[\textwidth][c]{%
      \begin{minipage}{.8\textwidth}
        \centering
        {%
          \Huge
          \bfseries
          A visual Brain-Computer Interface for gaze-free communication
        }
        \bigskip

        \MakeUppercase{%
          \Large
          Algorithms \& Applications
        }
        \bigskip

        {%
          \Large
          Arne  Van Den Kerchove
        }

      \end{minipage}%
    }

    \vfill
    \begin{minipage}[b]{.5\textwidth}
      Jury: \\
      Supervisor KU Leuven: \emph{prof. Marc Van Hulle} \\
      Supervisor University of Lille: \emph{prof. François Cabestaing} \\
      Chair examining committee: \emph{prof. John Creemers} \\
      Chair public defence: \emph{prof. Koen Poesen} \\
      Jury members: \\
      \emph{prof. Maarten De Vos} \\
      \emph{prof. Adalberto Simeone} \\
      \emph{prof. Andrea K\"ubler} \\
      \emph{prof. Fabien Lotte} \\
      \emph{prof. Reinhold Scherer}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.31\textwidth}\raggedright
        Dissertation presented in partial fulfillment of the requirements for
        the double degree of \emph{Doctor of Biomedical Sciences} (KU Leuven)
        and \emph{Doctor of Control Science and Signal Processing} (University
        of Lille)
    \end{minipage}
    \bigskip

    \centering\today

\end{titlepage}
\restoregeometry