### 2024-10-07 figures/wcble/smearing.tikz.tex

\sffamily
\sansmath
\begin{tikzpicture}   % First subplot (low jitter) positioned relative to center
    \begin{axis}[
        at={(0,0)}, % Adjust position to the left, closer to the center
        anchor=center,
        width=.45\textwidth, height=.309\textwidth,
        xmin=20, xmax=80,
        ymin=-0.1, ymax=1.1,
        axis lines=none, % Remove axes
        title={low jitter}, % Add title
        title style={yshift=-10pt, color=muteblack}, % Adjust title position to make it closer to the plot
        ]
        % Plot individual waveforms (low jitter) in darkgray
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-52)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-48)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-51)/5)^2)};
        \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-49)/5)^2)};

        % Plot average waveform (low jitter) in accent1 color, very thick
        \addplot[ultra thick,accent1,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};
    \end{axis}

   % Second subplot (high jitter) positioned relative to center
   \begin{axis}[
      at={(0.55\textwidth,0)}, % Adjust position to the left, closer to the center
       anchor=center,
        width=.45\textwidth, height=.309\textwidth,
       xmin=20, xmax=80,
       ymin=-0.1, ymax=1.1,
       axis lines=none, % Remove axes
       title={high jitter}, % Add title
       title style={yshift=-10pt, color=muteblack}, % Adjust title position to make it closer to the plot
       ]
       % Plot individual waveforms (high jitter) in darkgray
       \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-45)/5)^2)};
       \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-55)/5)^2)};
       \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-40)/5)^2)};
       \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-60)/5)^2)};
       \addplot[darkgray,domain=20:80,samples=100] {exp(-0.5*((x-50)/5)^2)};

       % Plot average waveform (high jitter) in accent1 color, very thick
       \addplot[ultra thick,accent1,domain=20:80,samples=100] {0.2*(exp(-0.5*((x-45)/5)^2) + exp(-0.5*((x-55)/5)^2) + exp(-0.5*((x-40)/5)^2) + exp(-0.5*((x-60)/5)^2) + exp(-0.5*((x-50)/5)^2))};
   \end{axis}

   % Unified legend positioned further below the center
   \node[anchor=center] at (.3\textwidth,-.2\textwidth) {\begin{tikzpicture}
       \begin{axis}[
           hide axis,
           xmin=0, xmax=0, ymin=0, ymax=0,
           legend style={
	   	draw=none, legend columns=2, /tikz/every even column/.append
	 style={column sep=2cm}
	 }
           ]
           \addlegendimage{darkgray,thick}
           \addlegendentry{Waveforms}
           \addlegendimage{accent1,ultra thick}
           \addlegendentry{Average}
       \end{axis}
   \end{tikzpicture}};
\end{tikzpicture}
%\vspace{-5cm}