### 2024-10-07 chapters/conclusion.tex
All text

\chapter{Conclusion \& recommendations}

\section{Contributions}

This work aimed to improve the performance of visual gaze-independent \acsp{bci}, in general
and applied to individuals with eye motor impairment by developing novel decoding
strategies and evaluating their effectiveness.
We addressed the limitations of current gaze-dependent \acp{bci} by proposing methods that
exploit covert \ac{vsa} and reduce the reliance on eye gaze.
The following key contributions were made:

\subsection{Developed \acs{erp} decoders}
We introduced a covariance estimator using adaptive shrinkage (STBF-shrunk) and an estimator
exploiting prior knowledge about the spatiotemporal nature of the EEG signal
(STBF-struct).
We compared these estimators with the original formulation of the
spatiotemporal (STBF-emp)
beamformer and a state-of-the-art Riemannian Geometry method (XDAWN+RG) in an off-line P3 detection task on
an existing dataset.
Our results show that the structured estimator results in higher accuracy in
general, and specifically when training data are sparsely available.
Results can be computed faster and with
substantially less memory usage.
Since these algorithms are not paradigm-specific, the conclusions can be
generalized to other ERP-based BCI settings.
These results have been published in~\textcite{VanDenKerchove2022}.

Next, \acf{bttda} was introduced as a tensor-based decoder that better captures the
multidimensional nature of \ac{erp} data.
By preserving the structure of neural data, \ac{bttda} can cope with noise and other challenges
arising from gaze-independent \acp{bci}, yielding robust results under certain conditions.
Results show that \ac{bttda} and its special sum-of-rank-one structured case
improved over \ac{hoda} and can reach state-of-the-art decoding performance for
\acp{erp}.
This work is submitted as~\textcite{Vandenkerchove2024b}.

Finally, we developed the \acf{wcble} decoder, specifically designed to address
the challenges posed by P3 latency jitter in covert \ac{vsa} settings.
Latency variability is a well-known issue in \ac{erp} decoding, particularly
when users engage in covert attention, where the timing of the P3 component
fluctuates significantly across trials.
Traditional decoders struggle in these settings due to the inconsistent timing
of brain responses, which can reduce the robustness and accuracy of
\ac{erp} decoding.
The \ac{wcble} decoder mitigates this issue by introducing an iterative process
to estimate and align \ac{erp} latencies across trials.
This is performed in such a way that discriminative power of the \ac{erp}
signal is preserved and enhanced.
The method has been published in~\textcite{VanDenKerchove2024}.
We took care to design a method that can be applied to improve decoding, where
unseen incoming test data is not yet known.
\Ac{wcble} was first tested on synthetic data to evaluate its
effectiveness under conditions of controlled latency variability and noise.
These results showed that is is robust to higher noise and jitter compared to
a non-iterative method in a latency estimation task.
It was also shown that decoding accuracy was higher in high jitter and noise
settings compared to the non-iterative method and a sate-of-the-art decoder.

\subsection{Gathered datasets}
The CVSA-ERP dataset consists of recordings of 15 healthy participants, mean age
$26.38\pm3.15$ years.
The dataset was presented in~\textcite{VanDenKerchove2024}.
The experiment in this dataset implemented a visual odball \ac{bci} with six
circular targets in a hexagonal layout.
The gaze fixation of participants was carefully controlled to dissociate the
visuospatial attention of the participant and their eye gaze.
This allows us to study effects of gaze-independence on \ac{erp} decoding.
The CVSA-ERP dataset gives us insight in the ERP dynamics in overt, covert and the
novel split VSA condition, and confirms our hypothesis that P3 jitter has a
significant impact on performance in covert and split VSA.
It also confirmed the effects of covert \ac{vsa} on \ac{erp} component
amplitude hold for split \ac{vsa}.

Additionally, we gathered data from 7 individuals with \ac{sspgi} with conditions such as
amyotrophic lateral sclerosis (\ac{als}), Friedreich's ataxia (\ac{fa}), and stroke.
These participants exhibited varying degrees of eye motor impairment, such as
involuntary eye movements, ophthalmoplegia, and gaze fixation fatigue.
The data provide invaluable real-world evidence on how gaze-independent
\acp{bci} perform in populations that experience eye motor difficulties.
This dataset allowed us to evaluate the proposed decoding strategies in clinical settings and
highlighted the practical challenges of implementing gaze-independent \acp{bci} in these user
groups.

\subsection{Investigated gaze-independent visual \acsp{bci}}

We have evaluated our proposed \ac{wcble} algorithm, its non-iterative
counterpart and state-of-the-art decoders on the CVSA-ERP dataset as well as on a publicly available
dataset~\cite{Aloise2012} that also contains the overt and covert VSA conditions.
We evaluated the BCI decoding performance in a single-trial classification experiment,
as well as in a target selection experiment reflecting BCI operation.
Performance was significantly different between decoders, but this
result was significantly dependent on the \ac{vsa} condition and the dataset.
While \ac{wcble} was outperformed by the state-of-the-art in overt \ac{vsa}
decoding, but increased covert \ac{vsa} decoding.

For the split attention conditions in the CVSA-ERP
dataset, \ac{wcble} yielded a significant improvement over CBLE and the
state-of-the-art method only in some cases.
These results were corroborated by analyzing selection accuracy, which showed
similar behavior for both datasets, except in overt
VSA, where accuracy was not harmed by the lower
single-trial selection performance of \ac{wcble}.

To further study the gaze-independent performance of these algorithms, transfer
learning between \ac{vsa} conditions is studied to simulate conditions where an
individual with gaze impairment can end up in different \ac{vsa} settings within
a \ac{bci}operation session due to their lack of proper motor control.
When trained and evaluated on overt VSA data, our proposed \ac{wcble} algorithm
results in a significant decrease in performance compared the state-of-the art,
consistent with the within-condition results.
For all other pairs of training and evaluation VSA condition, however,
\ac{wcble} was equal or significantly better.

Later case studies explored gaze-independent
\ac{bci} performance in individuals with \ac{sspi} and varying eye motor impairments.
Participants showed different levels of gaze control, with some using overt
\ac{vsa}, while others preferred covert \ac{vsa}, resting their gaze and
mentally attending to targets.

The \ac{wcble} decoder improved decoding accuracy covert
\ac{vsa} settings with cued central gaze fixation.
This is of interest for gaze-independent decoding, but not as optimal as if the
gaze fixation was uncued.
The latter was tested in the free \ac{vsa} settion.
Here, participants were generally more comfortable when allowed to use their
preferred gaze strategy.
Decoding performance in free \ac{vsa} was often comparable to overt \ac{vsa},
showing the system's flexibility.

Together, these results show that there is an interest in developing a new class of ERP-BCI
interfaces for patients that prefer to choose their own gaze strategy, so as to
avoid the effort of redirecting their eye gaze at different spatial locations
on the stimulation screen in manners that might be uncomfortable for them.
Coincidentally, advances in gaze-independent decoding also build towards a
solution for the Midas Touch Problem in visual \ac{bci}.
Here a \ac{bci} user sometimes accidentally selects a target while not wanting
to give any input.
Decoding of true intention independent of eye
gaze, with the option of gazing without paying visuospatial attention to a
stimulus, would counteract this.

\section{Current \& future work}
On the decoder development side, we have partially implemented some interesting
extensions of the proposed methods.
The Kronecker-structured beamformer can be generalized to the \ac{lda} case.
Furthermore, we obtained promising results when extending the single
Kronecker product covariance model to a sum of Kronecker product terms, since
the \ac{eeg} covariance is probably better expressed by such a
structure~\cite{Bijma2005}.

To overcome the limitations introduced by the presence of multiple \ac{erp}
components to latency estimation, we have also developed a multi-component
version of the \ac{wcble} algorithm.
This algorithm should be able to separate mixed \ac{erp} component clusters
based on their temporal coherence, in a similar maner to~\textcite{Ouyang2017}.
It could also yield favorable decoding results in a broader range of settings
than overt \ac{vsa} since it should be able to account for the presence of both
visual and attentional \ac{erp}components}.

In gaze-independence decoding, our current efforts focus on,
counterintuitively, integrating eye tracking into the decoding strategy.
Since we are able to explicitly discern overt, covert and split \ac{vsa} from
the \ac{erp}, and different decoders perform best in different settings, it
could be helpful to derive the current setting from the gaze position and the
\ac{erp} and propose the most suited decoder for a given data point.
select best classifier using eye tracking

We also wish to apply the tensor approach to gaze-inependent decoding.
The problems introduced by jitter could be accounted for as in \ac{wcble} by
alignment, but also by other methods that model the possible time shifts.
In ordering the data as a Hankel tensor\footnote{Adding an extra mode with
time-shifted copys of the temporal response per channel.}, we should have a
data representation that is more robust to jitter, at the cost of increased
dimensionality.
This increased dimensionality can than be countered by a tensor method like
\ac{hoda} or \ac{bttda}.

Finally, we wish to expand on experimentation with individuals with
\acp{sspgi}.
We aim to revisit the patients with an on-line experiment and comfort
assessment, to establish the usability of the proposed interface and to be able
to compare its \ac{itr}  to the state-of-the-art literature.

\section{Limitations and recommendations}

\subsection{Decoding: keep it linear; structure it}

Throughout this thesis, we've noticed that linear or multilinear decoding
methods consistentcly outperform non-linear methods, like SVMs or those based
in Riemannian Geometry.
Within these linear models, regularization (shrinkage, covariance structure,
tensor structure/rank) is of paramount importance, since neural data is
inherently multivariate and \ac{bci} calibration time is minimized.
Proper regularization should impose some specific structure on linear model,
ideally reflecting properties of the signal.
When faced with new problem and few data relative to dimensionality, it is a
good first instinct to pick a simple, restrictive structure.

Yet, the brain is a complex organ, hence neural data is not linear in origin and
these regularizing assumptions based on prior knowledge are probably only
superficially true.
When optimizing for performance, it will soon become clear that, given a more
flexible model, there are
settings where these assumptions can be broken to better describe the ongoing
interactions.
For example, spatiotemporal beamforming or \ac{lda} with Kronecker-structured covariance
improves performance in low sample sizes, but is outperformed by more flexible
methods, e.g.\ \ac{lda} using a Kronecker-sum covariance model or \ac{tlda}
if training sample size increases.
Similarly, \ac{bttda} is more flexible

This calls for the development of more flexible (multi-)linear models that still rely on some
structure, but which are combined with efficient model
selection, such as proposed by \ac{bttda}.
These models try to combine the best of both worlds.

%\subsection{Visual? gaze-independent \ac{bci}}
%
%While visual \acp{bci} are often praised for their easy, non-invasive
%applicability, they are also inherently unreliable, require wearing the
%\ac{eeg} cap and force continuous stimulation upon the user, all while forming
%a rather unintuitive method of communication for humans who are used to speech
%or motor commands.
%
%Currently, the group of users that would directly be helped with \emph{only} a
%gaze-independent visual \ac{bci} is not that large.
%Even if some eye motor impairment is the most suited strategy for a patient, it will probably be
%improved if a combination of eye tracking, residual motor control and other
%biosensors are used.
%
%
%\todo{https://www.nature.com/articles/s44222-024-00239-5}
%very small group of interest, but available right now, no implantation needed,
%
%true power in the future is probably focused on
%invasive, active paradigms if recording technology and surgical methods
%improve lowering the risk. orders of magnitude higher itr and more natural
%communication.
%will probably become acceptable if risks of implantation and longevity
%increase.
%characterized by publication date, but might indicate it is of interest to
%heavily focus this research now on applied 'vraagstukken' with patients
%
%Why limit ourselves to the visual~\ac{bci} decoding problem postulated in this
%thesis?
%
%
\subsection{Optimize user experience}
While our work with \ac{bci} users with \ac{sspgi} results in interesting
preliminary results, it is currently impossible to make claims about the experience in long term
home use of the proposed systems as assistive technology.
This work focused heavily on classification methods, and solving problem
arising from gaze-independence through decoding.
A natural metric to then evaluate proposed decoders is \ac{erp} classification
performance.

However, classification performance is only one aspect of what makes a \ac{bci}
desirable as assistive technology.
\Ac{itr} in on-line operation is a better metric, and the ultimate goal
is user satisfaction.
The latter can depends on performance, comfort and perception.
Classification performance is a convenient metric since a researcher can perform a single
data collection and iterate on the results offline.
However, this can be too constrictive when seeking the most suited interface
for a specific (group of users).

In the optimal approach, decoder development, paradigm choice and interface
design should working together, such as is advocated
by~\textcite{Pan2022} and~\textcite{Fouad2020}, by unifying the engineering and
the clinical side.
Performance and abilities are probably depend in such a degree on the user,
that off-line decoding performance in healthy controls is probably not a great
predictor of on-line decoding performance and \ac{itr} in individuals with
\ac{sspi}.
Incremental gains in performance in decoding for a specific paradigm might
also be outweighed by how well that paradigm is adapted to the user, factoring in,
skills and abilities.
Longitudinal research should follow several patient groups specifically with gaze
impairment and experiment with specific paradigms, interfaces and decoders for
a given \ac{bci} user.

\subsection{Work \emph{with} individuals with \acs{sspi}}

A logical point that arising from this is the question of how to setup an
effective research project that aims to optimize \ac{bci} user experience, in
the context of he topics presented in this work.
The patient-research involvement framework postulates that, ideally, individuals with
\ac{sspi} should be involved at every stage of the research project, from conceptualization
to evaluation.
A good rule of thumb is that research should solves problems \emph{with} the
users, not \emph{for} them, implying that research projects should start from
the need of the \ac{bci} user population.

Instead of choosing a research target and then moving on to decoder
development, data collection with healthy controls, patient studies and,
finally, verification in on-line operation, it could be more effective to start
the other way around.
In a mature field like visual oddball \ac{bci}, the research project can, after
a thorough literature review, immediately start with on-line experimentation
involving individuals with \ac{sspgi} and an adequately chosen, existing \ac{bci} system.
It should then become apparent where challenges for the envisioned user
(population) lie.
As an example, \textcite{FriedOken2020} share an overview of their experiences from this step,
specifically concerning eye motor impairment in visual \ac{bci}.
After that, hypotheses can be formed on how to mitigate these challenges
through interface optimization or choice taking into account the full picture
of the capabilities and skills of the users.
These hypotheses can then verified in experiments
with the users and/or a population of healthy controls.
Finally, the decoder can be optimized, only if it is strongly suspected that
effects will hold in on-line operation.

Unfortunately, this is also probably one of the least practical ways to set up
a 4-year research project, e.g.\ in the context of doctoral thesis.
This approach can only work if the proper facilities can be gathered immediately
from the start, like ethical approval and access to patients and on-line \acp{bci}
experimentation systems.
It is therefore important that research labs keep a working, long-term
collaboration with patient centers such that new projects can immediately be v
verified with population, and that interested users can compare multiple systems
and guide development.
The fact that a user has previous \ac{bci} experience is often seen as a confounding
factor in off-line studies.
When the research is in a phase that is heavily dependent on user
experimentation,
this is in fact a strength, sinche the user might help the researcher gain
insight into what works and doesn't work for them, and where research efforts
should be focused.

Similarly, there is value in maintaining a working, on-line in house \ac{bci}
system that can be iterated upon and adapted if necessary.
Experiments can be supported by advance calculations of hypothetical
\ac{itr}, such that performed on-line or off-line experiments reasonably reflect
real-world operation and no effort is wasted exploring settings that should not
reach satisfactory performance.
For example, why study a specific number of repetitions or \ac{isi}, if they
would not be practical to achieve a high \ac{itr}, but can still influence
decoding performance?

This top-down approach is probably not suited for all types of \ac{bci}
research.
Yet, in visual oddbal \ac{bci} where the challenge lies no longer in generating
a proof of concept but in translating the concept into a useful technology, it
might help avoid prematurely solving problems that might not necessarily be
encountered further down the road.
A top-down approach might improve the quality of study outcomes by enabling the
reporting of realistic, interpretable metrics,
and take into account the human aspect of working \emph{with} \ac{bci} users.

\section{Conclusion}