### 2024-10-07 chapters/funding.tex

\chapter*{Funding \& acknowledgments}
\todo{my grant}\todo{marc's projects}

Arne Van Den Kerchove is supported by the special research fund of the KU Leuven
(GPUDL/20/031).
prof. Marc Van Hulle is supported by research grants received from
the European Union’s Horizon Europe Marie Sklodowska-Curie Action program
(grant agreement No. 101118964), the European Union’s Horizon 2020 research and
innovation program (grant agreement No. 857375), the special research fund of
the KU Leuven (C24/18/098), the Belgian Fund for Scientific Research – Flanders
(G0A4118N, G0A4321N, G0C1522N), and the Hercules
Foundation (AKUL 043).

We thank the Flemish Supercomputer Center (VSC) and the High-Performance
Computing (HPC) center of the KU Leuven for allowing us to execute our
computational experiments on their systems.

We also thank,  Arno Libert,
Tjaša Mlinarič, Yide Li, Qiang Sun and Mani Mirsaeedi for their help in collecting data
for the works presented in \autoref{sec:covert-align} and
\autoref{sec:patients}.
and in general the colleagues at the KU Leuven Computational Neurosciences
group at KU Leuven and the BCI team at the University of Lille.
Furthermore, we thank the support teams at TRAINM Neuro Rehab Clinics Antwerp,
the UZ Leuven Neuromuscular Reference Renter and the Fondation Partage et Vie.
Finally, we wish to acknowledge dr.\ Axel Faes for his inspiration in conceptualizing
the work presented in~\autoref{sec:bttda}.

\todo{people at TRAINM, MAS and Kristl Claeys}
\todo{juliette meunier}