### 2024-10-07 fig.tikz.tex

\documentclass[tikz]{standalone}
\usepackage{dissertation}
\begin{document}
\input{figures/patients/fig_cross.pgf}
\end{document}