### 2024-10-08 chapters/wcble.tex

\chapter{\Acs{erp} latency estimation and alignment}
\label{sec:wcble}
\emph{Sections~\ref{sec:wcble/literature/contrib} and~\ref{sec:wcble/methods} were published
as part of~\textcite{VanDenKerchove2024}}

\section{Introduction}
\label{sec:wcble/intro}

\Ac{eeg}-based visual oddball \acp{bci} establish communication with paralyzed
individuals by decoding \acp{erp} obtained from time-modulated flashing
stimuli (see~\cref{fig:bci/loop}).
These \acp{erp} consist of multiple \ac{erp}-components, peaks, and troughs in
\ac{erp} waveform (see \cref{fig:bci/erp}).
These components each have their own distinct amplitude and latency.
\Ac{erp} component latency can reflect a wide range of cognitive,
neurophysiological, and clinical properties, that can be confounding factors in
the development of \acp{bci}~\cite{Luck2014}.
Latency is often assessed as the timing of an \ac{erp} component peak, obtained
after averaging over trials.
Latencies of non-averaged, single-trial \acp{erp} should capture even more
relevant information, yet they are very hard to measure due to the low \ac{snr} of
single-trial \acp{erp}.
Within-session variability of \ac{erp} latency, called \ac{erp} latency
\emph{jitter}, has been shown to influence \ac{bci}
accuracy~\cite{Thompson2012}.
This effect is especially prominent when bringing \ac{bci} development from the
lab to the clinical setting with patients in the target population.
\textcite{Zisk2021} showed that patients with \ac{als} have significantly
increased \ac{erp} jitter.
\textcite{Arico2014} showed that latency jitter negatively affects \ac{bci}
performance, and disproportionately so when the user is not directly gazing at
intended targets, such as is the case for individuals with \ac{sspgi}.
Therefore, properly accounting for \ac{erp} latency jitter can contribute to
the development of gaze-independent visual \acp{bci}.

\subsection{Neural origins of \acs{erp} latency jitter}
While the classic \ac{erp} model assumes a constant \ac{erp} response over trials,
research has shown that cortical evoked potentials show significant
trial-to-trial jitter~\cite{Truccolo2002}.
Experiments show that latency jitter in \acp{erp} is the result of the interaction of
the evoked activity with ongoing dynamics in the brain~\cite{Hasenstaub2007,
	Kisley1999, Curto2009, Arieli1996}.
According to the Firefly model of event-related activity~\cite{Burgess2012},
the latency of an evoked potential results from a combination of the modulation
of oscillating activity that is time-locked, but not necessarily phase-locked to stimulus onset,
and a delay caused by ongoing neurophysiological processes that are not necessarily related to the task
\cite{Stokes2016,Mouraux2008}.
If these processes are related to the \ac{bci} task at hand, such as attention
and visual processing, jitter could have a significant impact on \ac{erp}
decoding.

\subsection{Implications for \acs{erp} analysis}

The presence of latency jitter introduces multiple issues to \ac{erp} analysis, the
most prominent of which is the smearing effect, as illustrated
in~\cref{fig:wcble/smearing}. In \ac{erp} analysis, single-trial \acp{erp} are usually
averaged per condition to cope with the unfavorable \ac{snr}.
When averaging over multiple identical responses with jittered latencies, the shape
and amplitude of the average do not reflect the properties
of the original signal.
Because the amplitude of the average is lower, the \ac{snr} will also be negatively impacted.
More generally, the smearing effect leads to difficulties interpreting \ac{erp}
results and making inferences about amplitude or latency effects because these
are entangled when using averaging~\cite{Luck2014}.

\begin{figure}
  \centering
  \input{figures/wcble/smearing.tikz.tex}
  \caption[The smearing effect.]{The smearing effect. When obtaining an \ac{erp} template through
  averaging, high jitter causes a decrease in peak amplitude and a deformation
  of the shape of the waveform.}
  \label{fig:wcble/smearing}
\end{figure}

The smearing effect equally impacts the \ac{snr} of information captured by a
classifier's parameters when training with a procedure that does not account for
this jitter, and thus also affects its performance~\cite{Thompson2012}.
As an illustrative example, think of an \ac{lda} \ac{erp} classifier that uses
class averages over \ac{erp} epochs to construct its between-class scatter
matrix.
These per-class averages will be affected by the smearing effect and will be
less effective templates for \ac{erp} retrieval.
Similarly, the noise model in \ac{lda} assumes stationarity and subtracts
the class-wise averages from the signal to calculate
the within-class scatter matrix, which will then also be affected.
Hence, effective latency estimation and jitter compensation can contribute to higher
decoding performance.

\subsection{Benefits of latency estimation}
As we will see later, a common approach to counteract the smearing effect is to correct single-trial
\acp{erp} for jitter before averaging.
This can be done by estimating the latency of each
single-trial \ac{erp} and aligning them before averaging.
Latency estimation has multiple advantages:
\begin{enumerate*}[label={(\arabic*)}]
	\item Latency-based \ac{erp} alignment can help reveal the true shape and
	properties like peak amplitude and latency of \acp{erp} across
	experimental conditions. These are variables of interest for
	testing electrophysiological and behavioral hypotheses.
	Additionally, latency estimation allows for the extraction of
	other descriptors from a set of single-trial \acp{erp}, like
	amplitude or latency variance~\cite{Hultsch2004}, which can
	also be of interest as studied variables.
  For instance, in the work of \textcite{Saville2014}, the inter-trial latency
  variability itself is determined as a correlate of working memory performance.
	\item Counteracting the smearing effect will boost the signal-to-noise
	ratio of averaged \acp{erp} by increasing the amplitude of the
	\ac{erp} signal. This in turn can reveal smaller or more jittered \ac{erp}
  components which would otherwise be obscured or blurred due to desynchronization.
    These \emph{deblurred} templates can then be used for analysis of the \ac{erp}
	waveform, or as templates of the signal of interest in
	\acp{bci}~\cite{Arico2014}.
	%this enables decoders to take into account the relatively low amplitude
	%earlier components, leading to applications in settings where
	%information from other components than the P300 need to be exploited,
	%like covert visual attention paradigms~\cite{arico2014influence,
	%hardiansyah2020single}.\todo[inline]{rework this to latency features
	%because smaller visual components might not really be revealed
	%since they're already time-locked}
	\item Finally, the introduction of latency estimation and alignment generally
	improves \ac{bci} classification performance.
  Single-trial latencies can be used as features to improve
  classification~\cite{Hardiansyah2020}.
	Furthermore, compensating for inter-trial differences, especially when
	extended to inter-session or inter-subject differences, can aid in transfer learning or
	generalization across subjects and protocols~\cite{Iturrate2014}.
\end{enumerate*}

\section{Literature}

\subsection{Single-component approaches}
\label{sec:wcble/literature/single-comp}

The simplest method to estimate single-trial latency is
\emph{peak picking}.
Here, the latency of an \ac{erp} is determined as the
time-point of its maximum (or minimum) amplitude relative to stimulus
onset.
While straightforward, this method does not perform well in low
\ac{snr} conditions, unless combined with filtering.
Filtering can suppress the noise contaminating \ac{eeg} trials, lowering
the risk of picking a noise peak instead of a true peak of the \ac{erp}.
Several filtering approaches have been developed to be used in
conjunction with peak-picking, in the time domain~\cite{
  Carlton1980, McCarthy1981, Karjalainen1999, Nishida1999, Sparacino2002,
  Georgiadis2005, D’Avanzo2011}
and in the time-frequency domain~\cite{
  Quiroga2003, Wang2007, Hu2010}.
The aforementioned analysis and performance prediction method proposed
by \textcite{Arico2014} estimates the single-trial P3 latencies of attended
epochs by P3 peak-picking after filtering in the time-frequency domain and
corrects for jitter by temporally aligning all target epochs to the average
latency, i.e., by shifting the epochs in time such that the P3 peaks all fall
at the same time instance.
Filtering can
also be done in the spatial domain. The spatial filter can be constructed
using \ac{pca}~\cite{Ouyang2017}, \ac{ica}~\cite{Townsend1999, Jung2001,
	Milne2011}, or as a spatial LCMV beamformer~\cite{Treder2016}.

Template matching is generally preferred to peak picking.
The algorithm proposed by \textcite{Woody1967} in 1976 is a simple and elegant
latency estimation method that is still being used and is even
considered among the more performant techniques for
\acp{erp}~\cite{Thornton2008, Ouyang2017}.
It starts with an average \ac{erp} as a rough estimation of the aligned template
\ac{erp}, and iteratively refines this template by determining at each step the time point of maximum
cross-correlation of this template with all single-trial \acp{erp} and aligning them
on these time points to form the next template.
Similarly, cross-correlation template matching can be combined with filtering.
\textcite{Souloumiac2013} uses XDAWN to construct a spatial filter before
template matching the spatially filtered signal. \textcite{Iturrate2014} uses a
spatiotemporal filter. Instead of aligning the
latencies to obtain the template for the next iteration, weighted averaging can
also be used~\cite{Gasser1983}.

Despite good performance, the iterative scheme of Woody's algorithm has a risk of converging to a local
optimum.
To avoid this, a fitness function can be defined to jointly optimize
the set of single-trial \ac{erp} latencies, which can be optimized with a genetic algorithm~\cite{Pelo2018}.
Alternative methods are based on graph
optimization~\cite{Dimitriadis2018}, hidden process models~\cite{Kim2020}, or by fitting an \ac{erp} model with
maximum likelihood estimation~\cite{Gratton1989, Tuan1987, Moecks1988, Puce1994}.

\subsection{Multi-component approaches}
\label{sec:wcble/literature/multi-comp}
In the previous section, we have abstracted away the fact that \acp{erp} are
multi-component signals, which is not taken into account by the aforementioned
methods.
Multiple components can be time-locked to different events or neurophysiological
processes and therefore are not necessarily time-locked to each other. They
each can have different latency variabilities.
The presence of more than one component can hamper the
performance of some latency estimation algorithms, and it is often only possible to extract the latencies of the largest component present~\cite{Ouyang2017}. When aligning trials
to one component and averaging, this will introduce blurring in the other
components~\cite{Ouyang2020}.

Most of the algorithms above can be adapted to work on multiple \ac{erp} components
by carefully selecting peaks or pre-determined regions of interest.
\textcite{Hardiansyah2020} uses an \ac{snr} boosting method, applied to a specific
time-window for each \ac{erp} component to extract a latency per component.
Methods based on spatial decomposition filters like \ac{pca}, \ac{ica}, or XDAWN can be leveraged
to utilize multiple of their components as filters, resulting in multiple time
series where peak picking or Woody's algorithm can be applied. In general,
however, these algorithms lack an integrated approach to deal with multiple
\ac{erp} components. A clear evaluation of the performance of these
methods in multi-component settings is lacking.

Some algorithms separate components into stimulus- and response-locked component clusters based on their
latency distributions~\cite{Jung2001, Takeda2008, Zhang1998, Yin2009}.
These algorithms do, however, require a response
button trigger signal as input to determine the response time, which is
not always present or applicable in an \ac{erp} experiment, especially in
the case of \acp{bci}.
The RIDE algorithm~\cite{Ouyang2011, Ouyang2015, Wang2015, Ouyang2016, Ouyang2020}
is able to do this without entirely relying on the response time and can
separate stimulus-locked, response-locked, and multiple central component clusters.
Others, like spatial filtering methods that use decomposition methods to
determine the spatial filter, can be modified to separate
components~\cite{Ouyang2017}.

Finally, the smearing problem can also be overcome by algorithms that are not
based on \ac{erp} component latency estimation but use time-warping-based methods
to enhance the shape of the averaged template \ac{erp}, like Dynamic Time
Warping~\cite{Gupta1996, Wang2001, Zoumpoulaki2015}, Correlation-Optimized
Warping~\cite{Skov2006}, or Fast Variational
Alignment~\cite{Flotho2021}.
These methods work
fundamentally differently from the latency estimation methods, and cannot
directly be used to extract latency features.

\subsection{Contribution}
\label{sec:wcble/literature/contrib}

Most of these methods suffer from a common drawback: they cannot be used in a
decoding scheme to improve performance for incoming epochs with unknown
class labels that contain either a target or non-target \ac{erp} response.
They can be used offline on a set of labeled epochs for testing hypotheses
concerning latency and jitter, for aligning templates, or for \ac{bci} performance
prediction, but not for \ac{erp} classification.
While some of the aforementioned methods could be adapted to perform
classification tasks, few studies investigate how to exploit this latency estimation
for jitter-resistant decoding.
\textcite{Hardiansyah2020} incorporated single-trial latencies in
classification by peak-picking within a given
\ac{erp} time window, unaware of the class of the epoch under investigation.
The \ac{cble} algorithm introduced
by \textcite{Thompson2012} also explicitly applies latency estimation in a
decoding setting.
\textcite{Thompson2012} initially formulated \ac{cble} as an offline performance
prediction method.
Later, its output was successfully adapted to compensate for jitter to improve
decoder performance~\cite{Mowla2017, Zisk2022}.

Time-series classification algorithms~\cite{Abanda2019}
that are robust to jitter can be used in a decoding setting,
but, in general, have scarcely been applied to \ac{erp} decoding.
Data augmentation involving jittering the training
data~\cite{Krell2018, Zisk2022} and Riemannian Geometry methods using spatial
covariances as features~\cite{Aydarkhanov2020} have both been shown
to perform well in the presence of \ac{erp} jitter.
In this work, we opted to apply \ac{cble} because it has successfully been applied to
classify jittered \acs{erp}.
We adapt the \ac{cble} to an iterative method akin to Woody's scheme that can better
compensate for jitter, to improve covert VSA decoding performance.

\section{Methods}
\label{sec:wcble/methods}

\subsection{Classifier-based Latency Estimation}
\label{sec:wcble/methods/cble}

Consider a training set of $N$ EEG epochs with $C$ channels of $S$
samples $\{\mat{X}_n^\mathrm{train} \in \mathbb{R}^{C\times S}\}_{n=1}^N$
with corresponding training labels $\mathbf{l^\mathrm{train}} \in \{\mathrm{target},
	\textrm{non-target}\}_{n=1}^N$, and a
similar testing set of $M$ epochs $\{\mat{X}_m^\mathrm{test} \in
	\mathbb{R}^{C\times S}\}_{m=1}^M$.
We assume that the sampling period is $T$, i.e., that the sample with index $s$ is sampled at time $sT$.
In the following, we use the matrix slicing notation to denote row or column intervals extracted from a matrix.
For instance, $\mat{X}[:,s_1:s_2]$ denotes all columns of $\mat{X}$ with indices between $s_1$ (included) and $s_2$ (excluded).

\Ac{cble}, summarized in \cref{sec:wcble/app/cble-alg}, works by training a
\textit{first-stage} classifier $\mathcal{C}(\theta,f)$
defined within a time period $[s_1:s_2]$ with a set of parameters $\theta$ and a
decision function $f(\mat{X}[:,s_1:s_2],\theta) \rightarrow
y$ outputting a classification score $y\in\mathbb{R}$ for a given epoch $\mat{X}$,
such that
\begin{equation}
  \theta = \mathrm{train}_\mathcal{C}(\{\mat{X}_n^\mathrm{train}[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
  \label{eq:cble-train}
\end{equation}
Then,
$f$ can be applied to all (possibly overlapping) slices of length $s_2-s_1$ of
an epoch $\mat{X}$, resulting in a vector of score values
$\mathbf{y}=[y_1\,\ldots\,y_R]^T \in\mathbb{R}^R$ such that
\begin{equation}
  y_s = f(\mat{X}[:,s:s+(s_2-s_1)],\theta)\quad \forall s\in 1,\ldots,R
	\label{eq:score}
\end{equation}
with $R = S-(s_2-s_1)$.
To leverage \ac{cble} for \ac{erp} classification, the score vectors $\mathbf{y}$ can be
arranged in matrices $\mat{Y}^\mathrm{train}\in\mathbb{R}^{N\times R}$ and $\mat{Y}^\mathrm{test}\in\mathbb{R}^{M\times R}$.
These can be further classified by training a \textit{second-stage} classifier on
$\mat{Y}^\mathrm{train}$ and class labels $\mathbf{l}^\mathrm{train}$.
However, the resulting score-over-time vectors per epoch still suffer from jitter.
For classification, we follow the approach of \cite{Mowla2017}, using a
maximum-level hierarchical Daubechies-4 wavelet transform to reduce
dimensionality before classification with the second-stage classifier.
In the \ac{cble}-decoder, it is this wavelet transform that decreases the sensitivity
to latency differences, actively compensating for \ac{erp} latency jitter.

When using a simple spatiotemporal linear classifier as first-stage classifier,
\ac{cble} is equivalent to the first iteration of Woody's algorithm with the
spatiotemporal classifier weights as template.
\textcite{Thompson2012}, \textcite{Mowla2017}, and \textcite{Mowla2020} show
that \ac{cble} is relatively independent of the first-stage classifier for \ac{bci}
accuracy prediction and for \ac{erp} classification.
Therefore, we opt to use the variant of Linear Discriminant Analysis with block-Toeplitz
regularized covariance matrix (tLDA) proposed by \textcite{Sosulski2022}, as the first-stage
classifier and logistic regression as second stage.

\subsection{Robust latency features}
\label{sec:robust-latency}
The \ac{erp} decoding method based on \ac{cble} introduced by \textcite{Mowla2017} does
not make use of the extracted latencies, only passing score matrix $\mat{Y}$ on
to the second-stage classifier.
Furthermore, previous \ac{cble} works~\cite{Thompson2012, Mowla2020} only used these
latencies to correlate them with neurophysiological processes or to predict
decoder performance.
We noticed that, while \ac{cble} performance was unaffected, the classification
performance of our proposed method can be improved if the estimated
latencies are also made available as features to the second-stage classifier,
after a square transform for linear separability~\cite{Thompson2012}.

However, including these latency features gives rise to the following issue when
classifying unseen data.
The \ac{cble} latency estimate is defined only for target epochs as
$s_\mathrm{target}=\argmax_s{y_s}$.
This is the point in time where the target class reaches the largest separation
from the background noise and non-target class, indicating the target \ac{erp} is
most likely to occur here.
However, in a classifier test phase, it is not known a priori whether an
unseen epoch is a target or a non-target epoch.
This problem is solved by defining an estimated latency per class
$s_\mathrm{target}$ and $s_\mathrm{non-target}$ for every epoch,
regardless of its actual class.
The estimated class latencies can then be used as features for training and
testing the second-stage classifier.
This way, the latencies of the testing data can be presented to the second-stage
classifier without knowledge of the testing data class labels, making them
useful in decoding.

In a similar manner to the target latency, the non-target latency could be defined
as~$s_\mathrm{non-target}=\argmin_s{y_s}$.
However, this is problematic since it is not evident how to estimate
the latency of, e.g., a P3 \ac{erp} component for a non-target epoch, since the
non-target class is characterized by the absence of this component.
\footnote{Depending on the stimulation and experimental design, \ac{erp} components can be present in one experimental
	condition or class and missing in another, or appear in multiple classes
	at different amplitudes. In the latter case, it could be possible
	to estimate the latency of an \ac{erp} in the conditions where it appears, but if
	its amplitude is lower or negligible in some conditions, latency estimates
	will be less accurate, hence this case is still problematic.}.
In fact, $y$ can have multiple local minima or entirely lack distinct peaks for
non-targets, rendering the minimum estimate meaningless.

Instead, we opt for a more robust, probabilistic definition of class latencies.
This robust estimation method yields latencies that
\begin{enumerate*}[label=(\arabic*)]
  \item are more meaningful as input for the second-stage
    classifier, and
  \item lead to smoother convergence in our proposed iterative alignment scheme
    for \ac{wcble}, which heavily relies on exact latency estimation.
\end{enumerate*}
Assume classifier $\mathcal{C}(\theta,f,\Pr)$ now can also output a probability
per class $\Pr(l|
\mat{X}[:,s_1:s_2)], \theta)$ for a given epoch $\mat{X}$, a feature of many
common classifiers.
Analogous to equation~\ref{eq:score}, we can now write
\begin{equation}
\Pr(\mat{X},\theta,l,s) = \frac{1}{R}\Pr(\mat{X}[:,s:s+(s_2-s_1)],\theta,l)
  \quad \forall s\in 1,\ldots,R
	\label{eq:prob}
\end{equation}
The latency features assuming the epoch belongs to a given class
$l\in\{\textrm{target},\textrm{non-target}\}$ are then defined as the median of
the corresponding distributions
\begin{equation}
  s_l =\mathrm{median}\left[\Pr(s|\mat{X},\theta,l) \right]
  \label{eq:prob-latency}
\end{equation}
Note that $\Pr(s|\mat{X},\theta,\textrm{non-target}) =1-
  \Pr(s|\mat{X},\theta,\mathrm{target})$.
The median of the probability distribution over time is more robust to
outliers and noise than the maximum or minimum score.
For the non-target case, the median approach tends towards the center of a
near-uniform distribution, resulting in a more consistent latency estimate over
trials as compared to the minimum approach.

\subsection{\acl{wcble}}
To improve performance over \ac{cble}, we propose a new algorithm inspired both by
\ac{cble} and the aforementioned Woody iteration scheme termed \ac{wcble}.
Instead of using \ac{cble} to estimate the features of a second-stage classifier
directly, \ac{cble} latency estimation is used as a step in a Woody iteration scheme.
While the Woody algorithm iteratively enhances the \ac{snr} of an \ac{erp} template to
cross-correlate with the data, \ac{wcble} iteratively re-estimates the parameters of
the first-stage classifier.
To improve convergence and perform well in a classification setting, \ac{wcble}
aligns both targets and non-targets to their corresponding estimated latencies.

The \ac{wcble} algorithm is presented in \cref{sec:wcble/app/wcble-alg}.
Its training phase is visualized in~\cref{fig:diagram-train}.
The initial training epochs $\{\mat{X}_n^{(1)}\}_{n=1}^N$ are set to $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N$.
At every iteration, classifier $\mathcal{C}$ is trained like in \ac{cble}:
\begin{equation}
  \theta^{(i)} =
  \mathrm{train}_\mathcal{C}(\{\mat{X}^{(i-1)}_n[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
\end{equation}
Next, latency $s_{l_n}^{(i)}$ is determined for every epoch $\mat{X}^{(i)}$ corresponding
to its class label $l_n$ using \cref{eq:prob-latency}.
Finally, the training epochs $\mat{X}^{(i+1)}$ for the next iteration are determined by aligning
each original training epoch to the latency $s_{l_n}^{(i)}$ corresponding to its respective class
label.
\begin{equation}
  \mat{X}^{(i+1)}_n = \mathrm{align}(\mat{X}_n^\mathrm{train}, s_{l_n}^{(i)}) \quad \forall n=1,\ldots,N
\end{equation}
Aligning is performed by shifting and zero-padding the signal to the right if
the latency is negative relative to the time window onset, and to the left if
positive, by the difference between the latency and the window onset.
\fullpagefig{%
  \includegraphics[width=\textwidth]{figures/wcble/figure1.pdf}
}{%
  Schematic representation of the \acs{wcble} training phase.
}{%
  Schematic representation of the \ac{wcble} training phase.
  (1) The first-stage spatiotemporal binary classifier is trained on a set of
  epochs.
  (2) It is then applied to time-shifted copies of the epochs to obtain scores
  and class probabilities over time.
  (3) The medians of these probability distributions are assumed as the class
  latencies.
  (4) The epochs are aligned to their corresponding class latencies by
  shifting in time such that all latencies fall at the same moment.
  (5) The spatiotemporal classifier is then retrained on the aligned epochs
  for the next iteration.
  (6) After the iterative process halts, the scores and latencies obtained
  from the last iteration are used to train the second-stage
  classifier.
}{%
  fig:diagram-train
}
The process halts after a fixed number of iterations or when the estimated set
of latencies has been encountered before, indicating it ended up in a loop.
In the end, the procedure should result in enhanced classifier parameters $\theta^*$,
closer to those when there would be no jitter between epochs.
Note that using the median approach for robust latency estimation results in a
smoother yet longer convergence process compared to the maximum/minimum approach.
We can then apply the classifier with enhanced parameters $\theta^*$ in a \ac{cble}
manner to unseen epochs as illustrated in \cref{fig:diagram-eval} to obtain a
vector of scores over time as in~\cref{sec:cble} and the estimated latencies as
in~\cref{sec:robust-latency}.
\fullpagefig{
  \includegraphics[width=\textwidth]{figures/wcble/figure2.pdf}

}{%
  Schematic representation of the \acs{wcble} test phase.
}{%
  Schematic representation of the Woody Classifier-Based Latency
  Estimation test phase. (1) The
	first-stage spatiotemporal binary classifier obtained from the training phase is
	applied to time-shifted copies of the epochs to obtain scores and class probabilities
	over time. (2) The medians of these probability distributions are assumed
	as the new class latencies. (3) The scores and class latencies are input to the
  trained second-stage classifier, which predicts the label of the epochs.
}{%
  fig:diagram-eval
}
\section{Results}
\label{sec:wcble/results}
Since it is hard to obtain ground truth measures of \ac{erp} latencies, we
first evaluate our approach with synthetic data.
In \cref{sec:covert-align}, the proposed algorithm will be applied to real
\ac{eeg} data.

\subsection{Synthetic data generation}
To verify our approach, we simulated synthetic 16 channel \ac{eeg} data with standard
10-20 electrode positions using the \texttt{simulate\_evoked} method of the
MNE (version 1.8)~\cite{Gramfort2013} software package.
\Acp{erp} were generated by projecting a sine wave pulse source time course in a dipole in the
left hemisphere to the scalp electrodes using a boundary element method forward
model~\cite{Mosher1999} and MNE's \texttt{fsaverage} source space and anatomy.
Pink temporal noise is generated by passing Gaussian noise through an infinite
impulse response filter with transfer function
$\frac{B(z)}{A(z)} = \frac{1}{1 -1z^{-1}+0.15z^{-2}}$.
Pink spatial noise is also added, using a noise covariance matrix constructed from the
cosine distances between electrode positions and scaling it by the inverse of
the resulting spatial covariance.
We refer to \textcite{Gramfort2014} for implementation details.

The source time course is defined by the following function:
\begin{equation}
  s(t,l) =
  \begin{cases}
    a\sin\left(2\pi f\left(t-l\right)\right) & \text{if  $l-\frac{1}{2f} < t < l-\frac{1}{2f}$} \\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
over time $t$, with latency $l$, amplitude $a=\num{1e{-7}}$V, and frequency
$f=4$.

100 target epochs containing the evoked potential and noise and 100
non-target epochs containing only noise were simulated.
The target epochs were jittered by setting the latency offset $l$
of each target epoch to a random latency drawn from a normal distribution with 0 mean and standard
deviations of respectively $\sigma=0.1,0.2,0.3$s.
The noise was scaled at 32 different noise levels, with the \ac{snr} of a
target epoch ranging from 0 to -31dB, with
\begin{equation}
  \text{SNR} = 10\log_{10}\frac{\text{var}s(t,0)}{\text{var}\left(\text{noise}\right)}
\end{equation}

A sample of the simulated evoked data is displayed in \cref{fig:wcble/sim-sine}.
\fullpagefig{%
  \sffamily
  \sansmath
  \inputpgf{figures/wcble}{simulated-sine.pgf}
}{%
  Simulated target and non-target evoked data.
}{%
  Simulated target (bottom half of epochs) and non-target (top half of epochs) evoked data
  and their average (red: target, yellow: non-target, shaded area: standard
  deviation).
  Targets contain a sine wave pulse jittered by $\sigma$. 32 EEG channels with pink
  spatiotemporal noise were simulated using a forward head model, and channel C3 is
  plotted here.
}{fig:wcble/sim-sine}

\subsection{Latency estimation}
We compared our proposed \ac{wcble} algorithm to \ac{cble} on a latency
estimation task using the Pearson correlation coefficient $\rho$ at different
jitter and \ac{snr} levels.
For both \ac{cble} and \ac{wcble}, target latencies were extracted in a
10-fold cross-validation scheme using the robust median latency method after
fitting the model.
\Ac{wcble}'s maximum iteration number was set to 64.
Results are presented in \cref{fig:wcble/results/latency}.
A 95\% confidence interval on each measure was obtained through bootstrapping
with 1000 permutations.

\begin{figure}
  \sffamily\sansmath
  \input{figures/wcble/fig_latency.pgf}
  \caption[Synthetic latency estimation results]{Correctness of estimated
  latency for varying jitter ($\sigma$) and \ac{snr} introduced in synthetic evoked target
  data, measured as Pearson correlation coefficient $\rho$ between
  estimated and ground truth latencies.
  \Ac{wcble} estimates latencies more accurately in the presence of noise and
  high jitter.}
  \label{fig:wcble/results/latency}
\end{figure}

As the \ac{snr} decreases, both \ac{cble} and \ac{wcble} can accurately
estimate evoked potential latencies up to a given threshold. After this
threshold, correlation between estimated and ground-truth latencies will start
decreasing.
However, we notice that this threshold occurs later for \ac{wcble} than for
\ac{cble} for all evaluated jitter levels, indicating that \ac{wcble} is more
robust in the presence of pink spatiotemporal noise.
At around -28dB, \ac{wcble} does not significantly outperform \ac{cble}
anymore, and both methods get overtaken by the strong presence of noise.

\subsection{Classification}

Analogous to the previous experiment, 10-fold cross-validated classification
accuracy between target and non-target epochs was assessed.
Here, \ac{wcble} was compared to \ac{cble}, but both also to their base classifier
\ac{tlda}.
\Ac{cble} and \ac{wcble} classifiers were implemented as described
in \cref{sec:wcble/methods}, with both the wavelet-transformed score time series
and square-transformed latencies as input to a second-stage logistic regression
classifier with $L_2$-norm regularization ($C=0.2$).

\begin{figure}
  \sffamily\sansmath
  \input{figures/wcble/fig_accuracy.pgf}
  \caption[Synthetic decoding results]{Decoding performance, measured as binary
  classification accuracy, for \ac{cble}, \ac{wcble}, and their first stage
  classifier \ac{tlda} for varying introduced jitter ($\sigma$) and \ac{snr} in
  synthetic evoked target and non-target data.
  \Ac{wcble} decoding performance is robust to more noise and jitter than the other methods.}
  \label{fig:wcble/results/accuracy}
\end{figure}

\Cref{fig:wcble/results/accuracy} shows that the classification accuracy of both \ac{tlda} and \ac{cble} decreases with
$\sigma$ and \ac{snr}, but \ac{cble} outperforms \ac{tlda}.
This difference is particularly large when $\sigma$ is higher ($\sigma=0.2$
and $\sigma=0.3$) and \ac{snr} is not yet high.
\Ac{wcble} always scored better than \ac{cble} and \ac{tlda}, except for
$\sigma=0.1$ and $\text{\ac{snr}} < 27 \text{dB}$, at which point both \ac{cble}
and \ac{wcble} performed at chance level.
When considering the accuracy of \ac{wcble}, we see that it shows more of a
plateau as \ac{snr} increases.

\section{Discussion \& conclusion}

In this work, we introduced a new \ac{erp} latency estimation and decoding
method that applies a spatiotemporal classification iteratively along the time
dimension of the data, refining the dataset at each iteration by aligning
trials.
The algorithm combines \acf{cble} with Woody's template matching iteration
scheme and is therefore named \acf{wcble}.
Our method attempts to model the spatial and temporal activation of
each \ac{erp} component, as well as its latency distribution.
Through the first-stage classifier, it takes into account the
spatial and temporal structure of \ac{eeg} background noise and the trials from the
classes to discriminate.

The key finding of the simulation study presented
in \cref{sec:wcble/results} is that \ac{wcble} is robust to higher evoked
potential jitter and lower \ac{snr} than \ac{cble}.
It could therefore be a better candidate to apply in \ac{erp} analysis and
decoding in those cases where jittered \ac{erp} data is of interest.
At a certain noise threshold, both methods begin to fail, but \ac{wcble}
delivers higher accuracy at intermediate \ac{snr} levels than \ac{cble}, likely due to the iterative alignment process.

In general, noise beyond a certain level will overwhelm both \ac{cble} and
\ac{wcble} performance, making it impossible to accurately estimate latencies
or perform meaningful classification.
When latencies cannot be properly estimated, aligning epochs using \ac{wcble} does not help to improve classification.

As shown, the proposed method is suitable for \ac{bci} decoding settings
as well as single-trial \ac{erp} latency analyses.
In the case where there is only one data class present, the first-stage
classifier can be replaced by a spatiotemporal filter.
Instead of outputting a classification score, this filter should now output a
metric representing the presence of a template response in the signal.
A suitable candidate is the spatiotemporal
\ac{lcmv}-beamformer~\cite{VanVliet2015}, described in
\cref{sec:stbf-struct}.
The activation pattern can be set to the average of the set of trials to
analyze, and it can be refined at each \ac{wcble} iteration by taking the
average of aligned trials.

Finally, we argue that \ac{tlda} is a suitable first-stage classifier.
Firstly, imposing a Toeplitz-covariance structure strongly regularizes the
problem~\cite{Sosulski2022, VanDenKerchove2022}, benefiting decoding performance.
Secondly, this method has a synergy with \ac{cble} since both make the same
assumption about the short-time stationarity of the \ac{eeg} background noise
within an epoch.
\Ac{cble} does not retrain the first-stage classifier for each time shift but
rather trains it once within the given window.
After training, the classifier parameters represent some information about the
expected \ac{erp} waveform and background noise.
By applying the trained classifier to different time shifts, it assumes this
\ac{erp} waveform can be shifted in time, but since the classifier's information about
the background noise was only obtained from the initial window, \ac{cble} assumes
its properties do not vary throughout the epoch.
The block-Toeplitz covariance structure of \ac{tlda} also assumes that the
background noise represented by this covariance after subtracting the class
averages is stationary within the epoch~\cite{Sosulski2022}.

The main limitation of the proposed method is its lack of capacity to handle
multi-component \ac{erp} data.
In real \ac{erp} analysis settings, multiple components are usually present,
and each component has a distinct contribution to class discriminability and
latency distribution.
A single-component method will yield issues in interpretability and convergence
in the presence of other components, as it can `lock-on' to a given component
cluster and flatten out the others.
This can be avoided, however, by the choice of a proper region of interest
isolating solely the component of interest.
Current work focuses on extending \ac{wcble} to a multi-component setting, on
the one hand to improve latency estimation and decoding performance on real
\ac{erp} data, and on the other hand to apply it as an \ac{erp} component
separation method in the fashion of \textcite{Ouyang2017}.

The proposed implementation theoretically supports processing multi-class
data, but the convergence and correctness of the \ac{wcble} solution is yet to
be properly studied for this case.