### 2024-10-09 chapters/conflict.tex

\chapter*{Conflict of interest \& funding}
The authors declare no conflict of interest.

Arne Van Den Kerchove is funded by a bilateral cooperation program between KU
Leuven (special research fund of the KU Leuven GPUDL/20/031) and the university
of Lille under the Global PhD Partnership program.
Prof. Marc Van Hulle is supported by research grants received from
the European Union’s Horizon Europe Marie Sklodowska-Curie Action program
(grant agreement No. 101118964), the European Union’s Horizon 2020 research and
innovation program (grant agreement No. 857375), the special research fund of
the KU Leuven (C24/18/098), the Belgian Fund for Scientific Research – Flanders
(G0A4118N, G0A4321N, G0C1522N), and the Hercules
Foundation (AKUL 043).