### 2024-10-09 chapters/contribution.tex

\chapter*{Personal contribution}
Arne Van Den Kerchove conceptualized the works and performed the analyses in
this thesis, under supervision and in consultation with prof. Hakim Si-Mohammed,
prof. Marc Van Hulle and prof. François Cabestaing.
Dr. Arno Libert and Dr. Benjamin Wittevrongel consulted on the analyses performed in
\cref{sec:stbf-struct}
Data collection was performed by Arne Van Den Kerchove for the dataset
presented in \cref{sec:covert-align} and by Arne Van Den Kerchove and
Juliette Meunier for the dataset presented in \cref{sec:patients}.
Participant recruitment at TRAINM was performed by Alixe Willemssens under
supervision of dr. Edward Schietekatte, at Fondation Partage et Vie by Marie de
Moura under supervision of prof. Etienne Allart and at the UZ Leuven Neuromuscular
Reference Center under supervision of prof. Kristl Claeys and prof. Philip Van Damme.
Prof. Etienne Allart provided clinical consultation.