### 2024-11-14 chapters/preface.tex

\chapter*{Preface}
%The mind works in mysterious ways\ldots\
%Let's try to hack it!
%
%I'd like to acknowledge

personal acknowledgments