### 2024-11-14 cover_kul.tex

\begin{titlepage}
    \sffamily
    \begin{minipage}[t]{.5\textwidth}
    \vspace{0pt}
    KU Leuven \\
    Biomedical Sciences Group \\
    Doctoral School of Biomedical Sciences \\
    Faculty of Medical Science \\
    Department of Neurosciences
    \bigskip

    University of Lille \\
    Doctoral School MADIS \\
    UMR CRIStAL \\

    \end{minipage}\hfill%
    \begin{minipage}[t]{.3\textwidth}
      \vspace{0pt}
      \includegraphics[width=\textwidth]{figures/kul_docschool.png}
      \smallskip

      \includegraphics[width=\textwidth]{figures/ulille.png}
    \end{minipage}

    \vfill

    \noindent\makebox[\textwidth][c]{%
      \begin{minipage}{.8\textwidth}
        \centering
        {%
          \Huge
          \bfseries
          A visual Brain-Computer Interface for gaze-free communication
        }
        \bigskip

        \MakeUppercase{%
          \Large
          Algorithms \& Applications
        }
        \bigskip

        {%
          \Large
          Arne  Van Den Kerchove
        }

      \end{minipage}%
    }

    \vfill
    \begin{minipage}[b]{.5\textwidth}
      \raggedright
      \small
      Jury: \\
      Co-supervisor KU Leuven: \emph{prof. Marc Van Hulle} \\
      Co-supervisor Univ. Lille: \emph{prof. François Cabestaing} \\
      Chair examining committee: \emph{prof. John Creemers} \\
      Chair public defense: \emph{prof. Koen Poesen} \\
      \emph{prof. Andrea K\"ubler} \\
      \emph{prof.\ Fabien Lotte} \\
      \emph{prof.\ Reinhold Scherer} \\
      \emph{prof.\ Maarten De Vos} \\
      \emph{prof.\ Adalberto Simeone}
    \end{minipage}\hfill%
    \begin{minipage}[b]{.31\textwidth}
        \raggedright
        \small
        Dissertation presented in partial fulfillment of the requirements for
        the double degree of \emph{Doctor of Biomedical Sciences} (KU Leuven)
        and \emph{Doctor of Control Science and Signal Processing} (University
        of Lille)
    \end{minipage}
    \bigskip

    \centering
    Defended December 16, 2024

\end{titlepage}