### 2024-11-17 figures/gaze_independence/lis_venn.tikz.tex

\sffamily
\begin{tikzpicture}

    % Draw the outer lightgray box with ultra thick darkgray line
    \draw[draw=darkgray, ultra thick] (-4,-1.5) rectangle (4,5.5);
    % Label for the outer box in the top left corner
    \node[anchor=north west, text=darkgray] at (-4,5.5) {\textbf{\acs{bci} users}};

    % Draw the outermost circle for SSPI with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent1] (0,2) circle (3);
    \node[text=accent1, text width=15em, anchor=center, align=center] at (0,4)
    {\small\textbf{severe speech and \\ physical impairment (SSPI)}};  % Centered bold text with line color

    % Draw the middle circle for SSPGI with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent2] (0,1) circle (2);
    \node[text=accent2, text width=10em, anchor=center, align=center] at
    (0,2) {\small\textbf{severe speech, \\ physical and gaze \\ impairment
    (SSPGI)}};  % Centered bold text with line color

    % Draw the innermost circle for LIS with ultra thick lines and bold text matching its line color
    \draw[ultra thick, accent3] (0,0) circle (1);
    \node[text=accent3, text width=5em, anchor=center, align=center] at (0,-.1)
    {\small\textbf{Locked-in \\ Syndrome \\ (LiS)}};  % Centered bold text with line color

\end{tikzpicture}