### 2024-11-26 figures/bci/erp_target.tikz.tex

\begin{tikzpicture}[trim axis left]
    \begin{axis}[
        width=\textwidth,
        height=0.6180\textwidth,
        xlabel={Time (ms)},
        ylabel={Potential ($\mu$V)},
        axis lines=middle,
        ymin=-5, ymax=10,
        xmin=-100, xmax=800,
        ytick={-5,0,5}, % Y-ticks at -5, 0, 5
        xtick={200,400,600}, % X-ticks at 200, 400, 600, 800
      	ylabel near ticks,
      	xlabel near ticks,
    ]
      % ERP waveform using cubic splines for smooth connection
      \addplot[
          smooth,
          tension=0.5,
          style={very thick, accent1}, % Set the line color to accent1
      ] coordinates {%
          (-100, 0)  % Starting point
          (0, 0)     % Baseline
          (50, 0)    % Pre-P1
          (80, 2)    % P1
          (115, -3)  % N1
          (160, 3)   % P2
          (220, 1)   % N2
          (400, 7)   % P3
          (650, 1)   % LPP (Late Positive Potential)
          (800, 1)   % End point
      };

      % Annotations for the components
      \node[anchor=south, color=muteblack] at (axis cs:80,2) {P1};
      \node[anchor=north, color=muteblack] at (axis cs:115,-3) {N1};
      \node[anchor=south, color=muteblack] at (axis cs:160,3) {P2};
      \node[anchor=west, color=muteblack] at (axis cs:220,1) {N2};
      \node[anchor=south, color=muteblack] at (axis cs:400,7) {P3};
      \node[anchor=south west, color=muteblack] at (axis cs:650,1) {LPP};

    \end{axis}
\end{tikzpicture}