### 2024-11-27 chapters/acknowledgment.tex

\chapter*{Acknowledgments}
We thank Arno Libert,
Tjaša Mlinarič, Yide Li, Qiang Sun and Mani Mirsaeedi for their help in collecting data
for the works presented in \cref{sec:covert-align} and
\cref{sec:patients}.
We would also like to thank the other colleagues at the KU Leuven Computational
Neurosciences group at KU Leuven and the BCI team at the University of Lille,
for their general support.
Furthermore, we thank the teams at TRAINM Neuro Rehab Clinics Antwerp,
the UZ Leuven Neuromuscular Reference Renter and the Fondation Partage et Vie
specialized care home.
We wish to acknowledge dr.\ Axel Faes for his inspiration in conceptualizing
the work presented in~\cref{sec:bttda}.
Finally, we thank the Flemish Supercomputer Center (VSC) and the High-Performance
Computing (HPC) center of the KU Leuven for allowing us to execute our
computational experiments on their systems.