### 2024-11-27 fig.tikz.tex

\documentclass[booklet]{dissertation}
\begin{document}
	\input{figures/bci/erp_target.tikz.tex}
\end{document}