### 2024-11-29 chapters/patients_table.tex
Layout conclusion

\sffamily
\small
\begin{tabular}{@{}llrrrlllrr@{}}
  \toprule
  \textbf{ID}  & \textbf{Diagnosis} & \textbf{Age} & \textbf{Sex} & \textbf{Hand.} &
  \textbf{Speech} & \textbf{Trach.} & \textbf{Communication} &
  \textbf{W} & \textbf{KB} \\ \midrule
  PA1 & bulbar-onset \acs{als} & 58  & M   & L     & anarthric  & no          & tablet                 & 3  & 4 \\
  PB1 & \acs{fa} & 41  & M   & L     & dysarthric & no          & verbal                 & 3  & 3 \\
  PB2 & \acs{fa} & 43  & F   & R     & dysarthric & no          & verbal                 & 3  & 3 \\
  PB4 & \acs{fa} & 48  & M   & R     & dysarthric & no          & verbal                 & 3  & 3 \\
  PC2 & brainstem stroke & 43  & M   & R     & anarthric  & yes         &  \makecell[l]{prompting\\+eye movement} & 2 & 4  \\
  PC3 & brainstem stroke & 43  & F   & R     & anarthric  & yes         & letterboard            & 2  & 3 \\
  PC4 & \makecell[l]{left cerebellar stroke \\ (trombosis of the  basilar \\ artery)} & 54  & M   & R     & anarthric  & yes & letterboard & 2 & 3 \\
  \bottomrule
\end{tabular}