### 2024-11-29 chapters/preface.tex

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

%I would like to extend my sincere gratitude to those who have guided me
%throughout this journey. Prof. Marc Van Hulle, for the many opportunities y
%
%members of the jury
%Liesel
%
%Everyone who helped bring this project to a successful end
%Juliette, etienne allart, people at fondation partage et vie, Neuromuscular reference center
%(Kristl Claeys), Alixe and dr. Edward Schietekatte at TRAINM
%
%check notes