### 2024-12-04 chapters/preface.tex

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

The past four and a half years have been a stimulating, instructive, and intense
journey of scientific and self-discovery. What started as a vague interest
in the applications of machine learning in the biomedical domain rapidly evolved
into a fascination with neural signal processing when I came into contact with
it via my master thesis.
Ultimately, this was supercharged by the very high science-fiction factor
of Brain-Computer Interfaces.
There is something about both the brain and computers, and
their beautiful complexity arising from simplicity, that I just can't
seem to let go of.

Yet, it is the tangible impact on people's lives of neurotechnology
which first drew me to the field.
The biggest advantage my PhD project had to offer was
its holistic, end-to-end nature, involving everything from innovation in signal
processing and machine learning under the hood to EEG
experimentation and, eventually, working with end-user patients. I have
always considered this PhD, first and foremost, a learning experience.
An experience that has provided me -- after a lot of trial and error -- with the
toolkit to, hopefully, continue making this impact in the future by
advancing biomedical technology.

For this, I am grateful to my thesis supervisors. The combination of the
network, resources, advice, and feedback you provided helped this project
succeed.
I would specifically like to thank prof. Marc Van Hulle for the many
opportunities you have given me, prof. François Cabestaing for your close,
active, and involved supervision, and prof. Hakim Si-Mohammed for the many
sharp-witted ideas and contributions you brought to the table.
I also want to thank the members of the jury, prof. Andrea Kübler, prof.
Fabien Lotte, prof. Reinhold Scherer, prof. Maarten De Vos, and prof.
Adalberto Simeone, for the time and effort you have spent evaluating
this thesis and for your insightful feedback.

Furthermore, I would like to explicitly thank everyone else who helped
in realizing the work presented in this thesis. Juliette, your
proactive attitude and multidisciplinary perspective were an immense help in
data collection. Without you, this thesis would probably be missing its
last chapter, and I hope you enjoy your well-deserved upcoming academic
career.
Working with some of our end-user study participants would also
not have been possible without the medical advice of prof. Etienne
Allart from the Neurorehabilitation Unit at CHU Lille, as well as
the others involved in recruiting and interacting with
these participants: Alixe and dr. Edward Schietekatte at TRAINM Neuro
Rehab Clinics, prof. Kristl Claeys and prof. Philip Van Damme from the
Neuromuscular Reference Center at UZ Leuven, Marie and the team at
Fondation Partage et Vie, and all colleagues who helped me in data
collection and processing. Most importantly, I am grateful to everyone
who participated in our study, and I am sincerely moved by the positive
reception among those of you who could benefit from it. Of course, I
would also like to thank all other students, colleagues, and friends who
submitted themselves to having their hair messed up by yet another EEG
experiment.

Since my thesis was a joint PhD between two labs, one at KU Leuven in
Belgium and one at the University of Lille in France, I got to
experience working at two different research institutions
This formed an invaluable enrichment of my career until now.
It also means I've had the amazing luck of having double the amount of wonderful
colleagues.

In Lille, I've very much appreciated working with Gaël. Your presence
never failed to lighten up the mood, and I wish you smooth sailings
in completing your own PhD. Next, I want to thank Jimmy for showing
me the way through Lille and its campus life, and for all the tea. I
also enjoyed discussing with José, Jean-Marc, and Marie-Hélène.
Even though I was not always physically
present in Lille, you've always made sure I felt part of the team.
Finally, I'm glad that the SIGMA team was around to keep me in touch with
my computer science roots.

In Leuven, I want to begin with a special thanks to those of you who
guided me at the start of my PhD: Benjamin, for introducing me to
the wonderful world of ERPs and linear decoding, and Axel and Arno for
powering through COVID with me. As for the rest of you lot -- Aline,
Anahit, Aurélie, Barbara, Bob, Elvira, Eva, Li Ang, Liuyin, Mani, Mansoureh,
Qiang, Tjaša, Valentina, and Yide -- you've made my days brighter in many ways,
and I could always count on you, be it for advice, support or help.
I also wish to thank the students and interns I was fortunate enough to
supervise for their dedication and great work, as well as the students at NeuroTech Leuven.
You never failed to impress me with creative ideas and excellent
execution. A final shoutout to everyone in and around the
Neurophysiology lab: Alessia, Astrid, Christophe, Inez, Jonathan,
Kenneth, Lara, Noa, Saba, Sara, Sofie, Stijn, Wouter, and all others.

On a more personal note, I probably wouldn't have gotten here if it wasn't
for the support (and proofreading) I have \emph{always} experienced
from my family and friends.
Cato, Loes, Mama, and Papa, I think you've provided about the
best nest I could imagine to grow up in. To Aderik, Joran and Siebe: thank you for
carrying me throughout my university studies.
To Hanne, Ingrid, Jonah, Joshua, Kiefer, Mathi, Vik, Ward and Yana, I have but one
thing to say: Snek \emph{is} love, Snek \emph{is} life.

Some say that whoever is mentioned first on a paper is the most
important, since they're supposed to have done all the work.
Without the last author on there, however, most scientific publications would
either never materialize or go unnoticed.
Therefore, I would like to mention my partner Liesel here at the end.
Liesel, thank you for everything you do and have done
for me. This does not only include the past few months and your help with
practical arrangements, proofreading, and putting up with me as I was
-- maybe a bit too much -- focused on work, but much, much more
than that.
I still cannot believe how lucky I am to have you by my side.

As you turn this page and start reading the first chapter, I will also
start a new chapter in my life. You, on the other hand, will dive into
some 200 pages of algebra, convoluted procedures, and tables with
numbers. Enjoy!