### 2024-12-04 cover_kul.tex

{%
\small
\parindent0pt
  \sffamily
  \begin{minipage}[t]{.5\textwidth}
  \vspace{0pt}
  KU Leuven \\
  Biomedical Sciences Group \\
  Doctoral School of Biomedical Sciences \\
  Faculty of Medical Science \\
  Department of Neurosciences
  \bigskip

  University of Lille \\
  Doctoral School MADIS \\
  UMR CRIStAL \\

  \end{minipage}\hfill%
  \begin{minipage}[t]{.3\textwidth}
    \vspace{0pt}
    \includegraphics[width=\textwidth]{figures/kul_docschool.png}
    \smallskip

    \includegraphics[width=\textwidth]{figures/ulille.png}
  \end{minipage}

  \vfill

  \makebox[\textwidth][c]{%
    \begin{minipage}{\textwidth}
      \centering
      {%
        \huge
        \bfseries
        A visual Brain-Computer Interface for gaze-free communication
      }
      \bigskip

      \MakeUppercase{%
        \Large
        Algorithms \& Applications
      }
      \bigskip

      {%
        \Large
        Arne  Van Den Kerchove
      }

    \end{minipage}%
  }

  \vfill
  \begin{minipage}[b]{.6\textwidth}
    \raggedright
    Jury: \\
    Co-supervisor KU Leuven: \emph{prof.\ Marc Van Hulle} \\
    Co-supervisor Univ. Lille: \emph{prof.\ François Cabestaing} \\
    Chair examining committee: \emph{prof.\ John Creemers} \\
    Chair public defense: \emph{prof.\ Koen Poesen} \\
    \emph{Prof.\ Andrea K\"ubler} \\
    \emph{Prof.\ Fabien Lotte} \\
    \emph{Prof.\ Reinhold Scherer} \\
    \emph{Prof.\ Maarten De Vos} \\
    \emph{Prof.\ Adalberto Simeone}
  \end{minipage}\hfill%
  \begin{minipage}[b]{.31\textwidth}
      \raggedright
      Dissertation presented in partial fulfillment of the requirements for
      the double degree of \emph{Doctor of Biomedical Sciences} (KU Leuven)
      and \emph{Doctor of Control Science and Signal Processing} (University
      of Lille)
  \end{minipage}
  \bigskip

  \centering
  \hfill Defended December 16, 2024\hfill
}